Agile Monitoring of Evolving Web Services: For the Little Man

### 2018-04-04 discussion.tex
done or the day
cleaned up the performance testing part

>>>>

- time as measured in days and commits are two sides of the same coin.

- currently the system supports only one grouping. this need not be so, and there are many scenarios where one could envision multiple groupins in parallel.

- threats to the validity of the constant-load performance testing. performance regressions? 
- the endpoint where the unit-testing results can be uploaded is prone to being tampered with by Bob who would want to upload junk information to mess up with the dashboard. to solve this, a UUID can also be added that is generated by the server, and thus prevents anybody else but the rightful owner of the dashboard

- user management. although not explained in the paper, the dashboard also comes with a user management system. and two classes of users. admins and guests.

### 2018-04-04 evolution.tex
done or the day
cleaned up the performance testing part

%!TEX root=restructured.tex

  \section{Evolution monitoring }

@@

      dashboard.config.git = 'path/to/.git'

@@

    \begin{figure}[h!]
      \centering
      \includegraphics[width=\linewidth]{utilization-evolution}
      \caption{The Evolution of Endpoint Utilization Across System Versions}
      \label{fig:mv-util}
    \end{figure}

@@

    \begin{figure}[h!]
      \centering
      \includegraphics[width=0.9\linewidth]{translation_endpoint_evolution_}
      \caption{The Performance Evolution of the \epTranslations endpoint}
      \label{fig:tee}
    \end{figure}

@@

\begin{figure}[h!]
  \centering
  \includegraphics[width=0.7\linewidth]{time_per_user_per_version}
  \caption{This perspective shows that the evolution of response times for individual users (horizontal lines) across versions (the x-axis)}
  \label{fig:tuv}
\end{figure}

The colors represent users. The figure shows average performance varying  across users and versions with no clear trend: this is probably because varying user workload (i.e. number of sources to which the user is registered) is the reason for the variation in response times. \ins{one can see that for user 1 performance degrades over the }

>>>>

%!TEX root=paper.tex

  \section{LOC\#3: Monitoring Evolution}

@@

# LOC #3: provide the dashboard with 
# information on where to find the 
# git information 
dashboard.config.git = 'path/to/.git'

@@

    \begin{figure}[h!]
      \centering
      \includegraphics[width=1\linewidth]{utilization-evolution}
      \caption{The Evolution of Endpoint Utilization Across System Versions}
      \label{fig:mv-util}
    \end{figure}

@@

    \begin{figure}[h!]
      \centering
      \includegraphics[width=\linewidth]{translation_endpoint_evolution_}
      \caption{The Performance Evolution of the \epTranslations endpoint}
      \label{fig:tee}
    \end{figure}

@@

\begin{figure}[h!]
  \centering
  \includegraphics[width=0.97\linewidth]{time_per_user_per_version}
  \caption{This perspective shows that the evolution of response times for individual users (horizontal lines) across versions (the x-axis) for a given endpoint}
  \label{fig:tuv}
\end{figure}

The colors represent users. The figure shows average performance varying  across users and versions with no clear trend: this is probably because varying user workload (i.e. number of sources to which the user is registered) is the reason for the variation in response times. \ins{one can see that for user 1 performance degrades over the versions. Given that for other users the performance does not degrade in the same way, it is safe to assume that the problem might lay elsewhere.}

### 2018-04-04 grouping.tex
done or the day
cleaned up the performance testing part

For service endpoints which run computations in real time, the maintainer of a system might want to understand the endpoint performance on a per-user basis, especially for situations where the system response time is a function of some individual user load\footnote{E.g. in GMail some users have two emails while other have twenty thousand and this induces different response times for different users}.
\todo{remove the configuration part and rephrase the following, the discussion has been partially subsumed by Section~\ref{sec:tool}.}

@@

# attaching a group by function 
dashboard.config.group_by = 'User',
	lambda: Session.find(flask.request).userid

@@

@@

\todo{consider losing one or both of the figures --- or at least updating them}

@@

@@

\begin{figure}[h!]
  \centering
  \includegraphics[width=.8\linewidth]{time_per_user}
  \caption{The \epFeedItems shows a very high variability across users}
  \label{fig:tpu}
\end{figure}

>>>>

For service endpoints which run computations in real time, the maintainer of a system might want to understand the endpoint performance on a per-user basis, especially for situations where the system response time is a function of some individual user load\footnote{E.g. in GMail some users have two emails while other have twenty thousand and this induces different response times for different users}.

@@

# LOC #2: configure the dashboard
# to group requests by the user id
dashboard.config.group_by = 'User',
	lambda: Session.find(flask.request).userid

@@

Explain the code: the group by is assigned a tuple:

\begin{itemize}
	\item 'User' stands for the name of the grouping
	\item the ``lambda'' stands for a function that takes as argument
	the global flask.request. Every application must have a way of
	associating a user with a request. Or an app with a request. 
	A user is usually retrieved from some session variable. 
\end{itemize}

@@

@@

\begin{figure}[h!]
  \centering
  \includegraphics[width=.6\linewidth]{time_per_user}
  \caption{The \epFeedItems shows a very high variability across users}
  \label{fig:tpu}
\end{figure}

@@

### 2018-04-04 introductory.tex
done or the day
cleaned up the performance testing part

% \hfill mds
% Many contemporary programming languages are offering libraries, modules, or frameworks that facilitate the development of such architectures. 
Python is one of the most popular programming language choices for implementing the back-end of web applications. GitHub contains more than 500K open source Python projects and the Tiobe Index\footnote{TIOBE programming community index is a measure of popularity of programming languages, created and maintained by the TIOBE Company based in Eindhoven, the Netherlands} ranks Python as the 4th most popular programming language as of June 2016.

@@

\subsection*{NOTHING PREVENTS GENERALIZATION}
In this article we take as an example the Flask case study.
But there is nothing that prevents the generalization of 
our approach.

MAKE SURE THAT THIS IS CLEAR.

>>>>

% \hfill mds
% Many contemporary programming languages are offering libraries, modules, or frameworks that facilitate the development of such architectures. 
Python is one of the most popular programming language choices for implementing the back-end of web applications. GitHub contains more than 500K open source Python projects and the Tiobe Index\footnote{TIOBE programming community index is a measure of popularity of programming languages, created and maintained by the TIOBE Company based in Eindhoven, the Netherlands} ranks Python as the 4th most popular programming language as of June 2016. An analysis of  StackOverflow from September 2017 argues that {\em ``Python has a solid claim to being the fastest-growing major programming language''}\footnote{https://stackoverflow.blog/2017/09/06/incredible-growth-python/}.

@@

\subsection*{ENSURE CLARITY: NOTHING PREVENTS GENERALIZATION}
In this article we take as an example the Flask case study
and of the \tool as a dashboard. 
But there is nothing that prevents the generalization of 
our approach to other languages and technologies.

### 2018-04-04 testing.tex
done or the day
cleaned up the performance testing part

%!TEX root=restructured.tex

  \newpage
  \section{Preemptive monitoring}

  \ins{INTRO - LINKING WITH PREVIOUS PARAGRAPH: the previous evolutionary view crucial for 
  understanding the impact of software evolution on performance.
  However, just by observing this one can't be sure whether 
  performance changes because of the code or because of the payload.}

  \todo{the following might need a bit of rewrite}

  The concept of {\em preemptive monitoring} of the application performance by means of instrumenting integration unit testing as the synthetic load is similar to the idea of augmenting service monitoring with online testing~\cite{metzger2010proactive}, i.e.~testing service-based applications by using dedicated test input in parallel to its normal use and operation. The difference is that we take advantage of the capability of the CI framework (i.e. Travis in this case) to create an emulated ``live'' environment for integration testing purposes, and use unit testing as the dedicated test input in order to measure performance.

  For example, \Fref{fig:builds} is a screenshot from the dashboard showing the measured response times for 5 consequent builds, with 180 iterations of the unit tests executed in total across all endpoints. The outliers on the right of the figure are due to initial requests \ins{which must wait for a boot-up phase of the API}.

    \begin{figure}[h!]
        \centering
        \includegraphics[width=\columnwidth]{travis_builds}
        \caption{Response times in 5 consequent Travis builds in the \zee case study}
        \label{fig:builds}
      \end{figure}

@@

  \paragraph{Integration with Travis}
  \ml{added new section... also, this has to be moved after introducing the pre-emptive monitoring since it's about running tests. switched the order}

  \ins{The pre-emptive monitoring system} can be configured to work together with Continuous Integration (CI) frameworks like Travis\footnote{\url{https://travis-ci.org/}} that deal with automated integration testing.
  The developer needs to define the unit test folder of the project, the URL that Travis should use for reporting its results, and the number of iterations for each unit test should be executed in order to generate a load for the application.

  Each time a new build is created through Travis, the \tool automatically detects all available unit tests defined by the application developer and iterates through each one of them the number of times predefined by the developer while monitoring the response times for each request. The resulting measurements are persisted in a separate part of the tool database so as not to contaminate the ``live'' monitoring data. Beyond seeing the results of each Travis build in the \tool, this feature also allows for preemptive monitoring of the application, as we discuss in the following.

@@

\begin{figure}[h!]
  \centering
  \subfloat[The reported response time (in ms) per deployed version in the observation period]{\includegraphics[width=.5\columnwidth]{response_per_version_trunced_trend}}
  \quad
  \subfloat[The measured response time (in ms) using integration with Travis]{\includegraphics[width=.46\columnwidth]{travis_builds_no_outliers_trend}}
  \caption{Comparison of the response times per endpoint: actual production system versus preemptive monitoring data}
  \label{fig:preemptive}
\end{figure}

\begin{table}[h]
  \begin{tabular}{lll}
    \toprule
    Iteration & \bfseries Live (median) & \bfseries Travis (median)\\
    \midrule
    1 & 1349.41 & 764.87\\ 
    2 & 1466.13 & 992.87\\
    3 & 1256.65 & 760.87\\
    4 & 1266.42 & 813.89\\
    5 & 1080.68 & 303.4\\
    \bottomrule
  \end{tabular}
\end{table}

Pearson correlation $r(3)=.93, p=.02$

>>>>

%!TEX root=paper.tex

\newpage
  \section{Performance Regression Testing}

  What happens when the developer observes a performance degradation of a given API endpoint? How does he know whether the reason is the degradation of the performance of the code, the load on the machines, or maybe the change in the magnitude of the user data?

  \vspace{0.1cm}

  One way of solving the problem is to monitor the evolution of the performance of the endpoints measured with a constant load. Thus, if the performance decreases when an endpoint is tested with a constant load the problem must be the endpoint implementation; otherwise, it must be found elsewhere.

  \subsection*{Integrating with CI Servers}

  To monitor endpoint performance with a constant load \tool takes advantage of two best practices in software evolution: (1) and API must have unit tests for its endpoints, and (2) these unit tests will be run within a CI server.

  This raises the opportunity of piggybacking on the CI process to track the timing of these unit tests as a proxy of the endpoint performance evolution.

  \tool can be configured to work together with Continuous Integration (CI) frameworks like Travis\footnote{\url{https://travis-ci.org/}, one of the most popular CI solutions at the moment} that deal with automated integration testing. To do this one needs to simply add one extra line in the script that runs their continuous integration testing. In the case of Travis CI, a developer has to add the following line in `.travis.yml':

  \begin{lstlisting}[style=custompython]  
# LOC #4: to be added to the .travis.yml file
python -m flask_monitoringdashboard.collect
   --test_folder=./tests_zeeguu_api --times=5 
   --url=https://zeeguu.unibe.ch/api/dashboard

  \end{lstlisting}

  Each time a new build is created through Travis, the \tool automatically detects all available unit tests defined by the application developer (in (-{}-test\_folder)) and iterates through each one of them a number of times (-{}-times) while monitoring the response times for each test. The resulting measurements are uploaded to a specific endpoint that is created by the tool at the API url (-{}-url). The data uploaded to this endpoint is   persisted in a separate part of the tool database so as not to contaminate the ``live'' monitoring data.

      \begin{figure}[h!]
        \centering
        \includegraphics[width=0.45\columnwidth]{travis_builds}
        \caption{Response times in 5 consequent Travis builds}
        \label{fig:builds}
      \end{figure}

  \Fref{fig:builds} is a screenshot from the dashboard showing the measured response times for 5 consequent builds, with 180 iterations of the unit tests executed in total across all endpoints. The outliers on the right of the figure are due to initial requests \ins{which must wait for a boot-up phase of the API}.

  \subsection*{Preemptive Monitoring}
  The concept of {\em preemptive monitoring} of the application performance by means of instrumenting integration unit testing as the synthetic load is similar to the idea of augmenting service monitoring with online testing~\cite{metzger2010proactive}, i.e.~testing service-based applications by using dedicated test input in parallel to its normal use and operation. The difference is that we take advantage of the capability of the CI framework (i.e. Travis in this case) to create an emulated ``live'' environment for integration testing purposes, and use unit testing as the dedicated test input in order to measure performance.

  Figure \ref{fig:response_times_preemptive} shows one example of correlation between the actual endpoint response times across five versions of the system (above) and the corresponding testing times for the synthetic load in the same five versions (below). As suggested by the green line that unites the medians there seems to be a correlation between the two datasets as they are moving in the same.

      \begin{figure}[h!]
        \centering
        \includegraphics[width=0.75\columnwidth]{response_per_version_trunced_trend}

        \advance\leftskip-0.2cm
        \includegraphics[width=0.5\columnwidth]{travis_builds_no_outliers_trend}
        \caption{The reported response time (in ms) per deployed version in the observation period: 
        in the actual deployment (above) and using integration with Travis (below)}

        \label{fig:response_times_preemptive}
      \end{figure}

  Computing Pearson correlation ($r(3)=.93, p=.02$) between the median values of the two datasets (cf. Table \ref{tab:correlations}) shows that there is a correlation between the two.

    \begin{table}[h]

      \centering
      \begin{tabular}{lll}
        \toprule
        Iteration & \bfseries Live (median) & \bfseries Travis (median)\\
        \midrule
        1 & 1349.41 & 764.87\\ 
        2 & 1466.13 & 992.87\\
        3 & 1256.65 & 760.87\\
        4 & 1266.42 & 813.89\\
        5 & 1080.68 & 303.4\\
        \bottomrule

      \end{tabular}
      \caption{The measured response time (in ms) using integration with Travis}
      \label{tab:correlations}
    \end{table}

%   \begin{figure*}[hb!]
%   \centering
%   \subfloat[The reported response time (in ms) per deployed version in the observation period]{\includegraphics[width=.6\columnwidth]{response_per_version_trunced_trend}}
%   \quad
%   \subfloat[The measured response time (in ms) using integration with Travis]{\includegraphics[width=.56\columnwidth]{travis_builds_no_outliers_trend}}
%   \caption{Comparison of the response times per endpoint: actual production system versus preemptive monitoring data}
%   \label{fig:preemptive}
% \end{figure*}

@@

  \paragraph{Integration with Travis}

@@
